Testing Quasar Unification: I. Radiative Transfer in Clumpy Winds

### 2015-12-21 paper_mnras.tex
approaching submission ready

\maketitle
\begin{abstract} 
Various unification schemes have been proposed to interpret the complex phenomenology 
of quasars and luminous active galactic nuclei (AGN) in terms of  a simple axisymmetric 
picture involving a central black hole, an accretion disc and an associated outflow.
Here, we continue our tests of  this paradigm by comparing quasar spectra to 
synthetic spectra of simple biconical disc wind models, produced with our state-of-the-art 
Monte Carlo radiative transfer and photoionization code PYTHON.  
Previously, we have shown that we could produce synthetic spectra 
resembling those of observed broad absorption line (BAL) quasars 
for plausible wind models, but only if the X-ray luminosity was limited to about  
10$^{43}$~erg~s$^{-1}$. Here, we investigate the  degree to which
clumping of the outflow allows us to 
produce synthetic spectra in the rest-frame ultraviolet (UV) 
that have the characteristics of quasars with X-ray luminosities as 
large as 10$^{45}$ erg s$^{-1}$. 
We conclude, perhaps not surprisingly, that clumping does maintain the ionization 
state of the wind necessary for strong BAL features in the rest-frame UV at more 
realistic X-ray luminosities. We examine the X-ray properties of our fiducial clumped model
and find good agreement with existing X-ray samples of AGN and quasars.
In addition, the dense, X-ray heated wind produces strong recombination and collisionally 
excited line emission in, e.g., \la\ and \civline, to emerge at the low inclination, 
`Type 1 quasar-like' angles. At the highest inclinations, the synthetic spectra 
possess prominent \mg\ and \al\ BALs, the absorption features seen in LoBAL quasars. 
Despite these successes, we are unable to reproduce all the emission lines seen in
quasar spectra with the correct equivalent width 
ratios in any of the models in our initial simulation grid,
and we find an angular dependence of emission line equivalent width despite the 
remarkable similarities in the observed spectral properties of BAL and non-BAL quasars.
Overall, our work suggests that a biconical wind model
reproduces much of the qualitative behaviour expected from a unified model. 
However, we cannot yet provide quantitative matches with quasar properties at all 
viewing angles; whether disc winds can successfully unify quasars is therefore 
still an open question.
\end{abstract}

% \begin{keywords}
% galaxies: active; accretion, accretion discs; (galaxies:) quasars: emission lines;
% (galaxies:) quasars: absorption lines
% \end{keywords}

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities, $L$, represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure*}

\noindent
Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with clump densities 
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane, 
where $n_H$ is the local number density of H.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration results in
a wind that exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
With a clumped wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{erg~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles for the fiducial model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
SDSS J162901.63+453406.0, respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled to $1350$\AA
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*}

@@

Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999b,arav1999a}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show nonblack saturation.
Instead, we find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{hamann2008,surlan2012}.
Second, our wind does not have a significant scattering contribution 
along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc. We note the scattering contribution
from plasma in polar regions is significant in some `outflow-from-inflow'
simulations \citep{KP09, simproga2012}.

@@

We find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ\ and \nv\
lines. The improved treatment of recombination also results in a strong \la\ line. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To demonstrate this further,
we show line luminosities and monochromatic continuum luminosities
from the synthetic spectra in table~\ref{line_lums}. These are compared to
mean values from a subsample of the SDSS DR7 quasar catalog (Shen et al. 2011) 
with BH mass and Eddington fraction estimates similar to the fiducial model values 
(see caption). The spectra do not contain the strong 
C~\textsc{iii}]~1909\AA\, line seen in the quasar composite spectra,
because semi-forbidden (intercombination)
lines are not included in our modelling.

In figure~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites,
although velocity projection effects mean the Balmer lines are double peaked, 
as seen  in `disc emitter' systems \citep[e.g.][]{eracleous1994}.
The increased line equivalent widths (EWs) are due to a weakened continuum at high inclination from wind attenuation and 
disc foreshortening / limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
The high inclination viewing angle is a physically motivated quasar model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
% The critical density of the C~\textsc{iii}]~1909\AA\, line 
% is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
% outer portion of our wind. 
% We therefore expect a model with these 
% parameters to produce a C~\textsc{iii}]~1909\AA\ line
% with a proper treament. We note, however, the the presence of
% this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

@@

Despite a number of successes, 
there are a some properties of the synthetic spectra
that are at odds with the observations. First, we find that the ratios of the 
EW of the \la\ and \mgline\ lines
to the EW of \civline\ is much lower than in the composite spectra. 
This problem has also been seen in simpler photoionization models for the 
BLR \citep{netzer1990}.
It may be that a larger region of very dense ($n_e\sim10^{10}$cm$^{-3}$) 
material is needed, which could correspond to a disc atmosphere or 
`transition region' (see e.g. Murray et al. 1995; Knigge et al. 1997).
\nocite{knigge1998} 
While modest changes to geometry may permit this, the initial grid search 
did not find a parameter space in which the \la\ or \mg\ EWs
were significantly higher (see section~\ref{sec:param_sens}). 
Second, we find that EWs increase
with inclination 
(see figures~\ref{fig:uvspec} and \ref{fig:sed}; also figure~\ref{fig:lobal})
, to the extent that significantly denser
models can match the line EWs fairly well at low inclinations but will then
possess overly strong red wings to the BAL P-Cygni profiles at high inclinations
This angular dependence is expected from a model in which the BLR radiates
more isotropically than the disc, which in this case is foreshortened and limb darkened.
It appears to contradict observations, which show remarkably uniform emission
line properties in quasars and BALQSOs \citep{weymann1991,dipompeo2012b}. 
The angular distribution of the disc 
continuum and line emission
is clearly crucially important in determining the emergent broad 
line EWs, as suggested by, e.g., the analysis of \cite{risaliti2011}. 
We shall explore this question further in a future study.

@@

The parameter search presented here is by no means exhaustive, and
we may be limited by the specific parameterisation (from SV93) of the outflow 
kinematics we have used. Nevertheless, we suggest that the angular distribution
of both the line and continuum emission is perhaps the crucial 
aspect to understand. With this in mind, obtaining reliable orientation 
indicators appears to be a crucial observational task if we are to
further our understanding of BAL outflows 
and their connection, or lack thereof, to the 
broad line region.

@@

We have carried out MCRT simulations using a simple
prescription for a biconical disc wind, with
the aim of expanding on the work of H13 and assessing 
the viability of such a model for geometric unification of quasars.
We find the following main points:
\begin{enumerate}
\item We have introduced a simple treatment 
of clumping in our model, and found that it moderates the ionization state
sufficiently to allow for the 
formation of strong UV BALs while agreeing well with the X-ray
properties of luminous AGN and quasars. 
\smallskip
\item We find that a clumpy outflow model naturally 
reproduces the range of ionization states
expected in quasars, due to it's stratified density
and temperature structure. 
LoBAL line profiles are seen at a subset of viewing angles, and Fe~\textsc{ii}
absorption is seen at particularly high inclinations. 
\smallskip
\item We find that the higher X-ray luminosity 
causes a significant 
increase in the strength of the  emission
lines produced by the model. This is true both
of collisionally excited resonance lines (such as \civ, \nv)
and recombination lines (such as \la\ and \heiiline).
However, the equivalent width ratios of the emission lines do not match
observations, suggesting that a greater volume of dense ($n_e\sim10^{10}$~cm$^{-3}$)
material may be required.
\smallskip
\item The line EWs in the synthetic spectra increase with inclination.
BAL and non-BAL quasar composites have comparable EWs, so our model
fails to reproduce this behaviour.
 If the BLR emits fairly isotropically then for a 
foreshortened, limb-darkened accretion disc 
it is not possible to achieve line ratios at low inclinations 
that are comparable to those at high inclinations. 
We suggest that understanding the angular distribution of 
line and continuum emission is a crucial question for theoretical models.
% Furthermore, this is conclusion is independent of the assumed 
% BLR geometry and size.
\end{enumerate}
Our work confirms a number of expected outcomes from a geometric unification 
model, and suggests that a simple biconical geometry such as this can come close to 
explaining much of the  phenomenology of quasars. However, our conclusions pose 
some challenges to a picture in which BALQSOs are
explained by an {\em equatorial} wind rising from a classical thin disc, and suggest 
the angular distribution of emission is important to understand if this 
geometry is to be refuted or confirmed. We suggest that obtaining reliable 
observational orientation indicators and 
exploring a wider parameter space of outflow geometries in simulations
are obvious avenues for future work.

>>>>

\maketitle
\begin{abstract} 
Various unification schemes interpret the complex phenomenology of quasars and luminous active galactic nuclei (AGN) in terms of a simple picture involving a central black hole, an accretion disc and an associated outflow. Here, we continue our tests of this paradigm by comparing quasar spectra to synthetic spectra of biconical disc wind models, produced with our state-of-the-art Monte Carlo radiative transfer code. Previously, we have shown that we could produce synthetic spectra resembling those of observed broad absorption line (BAL) quasars, but only if the X-ray luminosity was limited to about 
10$^{43}$~erg~s$^{-1}$. Here, we show that clumping the outflow maintains the ionization state necessary for BAL features in the rest-frame UV at more realistic X-ray luminosities. Our fiducial model shows good agreement with X-ray properties of AGN and the dense, X-ray heated wind produces strong line emission in, e.g., \la\ and \civline at low inclinations. At high inclinations, the synthetic spectra possess prominent LoBAL features. Despite these successes, we cannot reproduce all the emission lines seen in quasar spectra with the correct equivalent width ratios, and we find an angular dependence of emission line equivalent width despite the remarkable similarities in the observed spectral properties of BAL and non-BAL quasars. Overall, our work suggests that a biconical wind can reproduce much of the qualitative behaviour expected from a unified model, but we cannot yet provide quantitative matches with quasar properties at all viewing angles. Whether disc winds can successfully unify quasars is therefore still an open question.
\end{abstract}

\begin{keywords}
galaxies: active; accretion, accretion discs; (galaxies:) quasars: emission lines;
(galaxies:) quasars: absorption lines; methods: numerical; radiative transfer.
\end{keywords}

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/link8.png}
\caption
{
Contour plots showing the logarithm of some important 
physical properties of the outflow. The spatial scales are
logarithmic and the $x$ and $z$ scales are not the same.
Symbols are defined in the text.
The solid black line marks a sphere at $1000~r_G$.
The dotted lines show the $72^\circ$ and $78^\circ$ sightlines 
to the centre of the system, and illustrate that different sightlines
intersect material of different ionization states.
The line luminosities, $L$, represent the luminosity of photons
escaping the Sobolev region for each line. These photons do not
necessarily escape to infinity.
}
\label{fig:wind}
\end{figure*}

\noindent
Figure~\ref{fig:wind} shows the physical properties of the wind.
The wind rises slowly from the disc at first, with densities within clumps
of $n_H \sim 10^{11}~\rm{cm^{-3}}$ close to the disc plane, 
where $n_H$ is the local number density of H.
The flow then accelerates over a scale length of $R_V=10^{19}~\rm{cm}$
up to a terminal velocity equal to the escape velocity at the streamline base
($\sim10,000~\rm{km~s^{-1}}$). This gradual acceleration results in
a wind that exhibits a stratified ionization structure, with low ionization material
in the base of the wind giving way to highly ionized plasma further out.
This is illustrated in figure~\ref{fig:wind} 
by the panels showing the ion fraction $F=n_j/n_{tot}$ of some important ions.
With a clumped wind, we are able to produce the range of ionization states observed
in quasars and BALQSOs, while adopting a realistic $2-10$ keV X-ray luminosity
of $L_{X}=10^{45}~\rm{erg~s^{-1}}$. Without clumping, this wind would be over-ionized 
to the extent that opacities in e.g., \civ\ would be entirely negligible (see H13).

@@

\begin{figure*}
\centering
\includegraphics[width=1.0\textwidth]{figures/uvspec_new.png}
\caption
{
Synthetic spectra at four viewing angles for the fiducial model. At 
$20^\circ$ and $60^\circ$ we show a comparison to an SDSS quasar composite
from Recihard et al. (2003). At $73^\circ$ and $76^\circ$ we show a comparison to
an {\sl HST} STIS spectrum of the high BALnicity BALQSO 
PG0946+301 (Arav et al. 2000), and an SDSS spectrum of the LoBAL quasar 
SDSS J162901.63+453406.0, respectively. The dotted line shows a disc
only continuum to show the effect of the outflow on the continuum level. 
All the spectra are scaled to the model flux at $2000$\AA, expect for the 
{\sl HST} STIS spectrum of PG0946+301, which is scaled to $1350$\AA\
due to the incomplete wavelength coverage.
% HST STIS spectrum of PG0946+301 (Arav et al. 2000)
}
\label{fig:uvspec}
\end{figure*}

@@

The absorption profiles seen in BALQSOs are often non-black, but saturated, 
with flat bases to the absorption troughs \citep{arav1999b,arav1999a}.
%Non-black saturation is observed in the absorption troughs of BALQSOs \citep{arav1999b,arav1999a}.
This is usually explained either as  partial covering of the continuum
source or by scattered contributions to the BAL troughs, necessarily
from an opacity source not co-spatial with the BAL forming region.
The scattered light explanation is supported by spectropolarimetry results
\citep{lamy2000}. Our spectra do not show non-black, saturated profiles.
We find black, saturated troughs at angles $i > 73^\circ$, and the BALs
are non-saturated at lower inclinations. The reasons for this are inherent 
in the construction of our model. 
First, the microclumping assumption does not allow for 
porosity in the wind, meaning that it does not naturally produce
a partial covering absorber. To allow this, an alternative approach
such as {\em macroclumping} would be required \citep[e.g.][]{hamann2008,surlan2012}.
Second, our wind does not have a significant scattering contribution 
along sightlines which do not pass through the BAL region,
meaning that any scattered component to the BAL troughs is absorbed by line opacity.
This suggests that either the scattering cross-section of the wind must
be increased (with higher mass loss rates or covering factors), or 
that an additional source of electron opacity is required, potentially
in a polar direction above the disc. We note the scattering contribution
from plasma in polar regions is significant in some `outflow-from-inflow'
simulations \citep{KP09, simproga2012}.

@@

Unlike H13, we now find significant collisionally excited line emission emerges
at low inclinations in the synthetic spectra, particular in the \civ\ and \nv\
lines. We also find a strong \la\ line and weak He~\textsc{ii}~$1640$\AA\ line
as a result of our improved treatment of recombination using macro-atoms. 
In the context of unification, this is a promising result, 
and shows that a biconical wind can produce significant 
emission at `quasar-like' angles. To demonstrate this further,
we show line luminosities and monochromatic continuum luminosities
from the synthetic spectra in table~\ref{line_lums}. These are compared to
mean values from a subsample of the SDSS DR7 quasar catalog (Shen et al. 2011) 
with BH mass and Eddington fraction estimates similar to the fiducial model values 
(see caption). The spectra do not contain the strong 
C~\textsc{iii}]~1909\AA\ line seen in the quasar composite spectra, 
but this is due to a limitation of our current treatment of C; semi-forbidden
(intercombination) lines are not included in our modelling.

In figure~\ref{fig:sed}, we show an $F_{\nu}$ spectrum with broader waveband coverage
that includes the optical, showing that our synthetic spectra 
also exhibit \ha\ and \hb\ emission. 
In this panel, we include a low inclination and 
also a very high inclination 
spectrum, which looks underneath the wind cone. This model shows 
strong line emission with very similar widths and line ratios to the quasar composites, and
the Balmer lines are double peaked, due to velocity projection effects.  
Such double-peaked lines are seen in so-called `disc emitter' systems 
\citep[e.g.][]{eracleous1994} but not the majority of AGN.     
The line equivalent widths (EWs) increase at high inclination
due to a weakened continuum from wind attenuation, 
disc foreshortening and limb darkening. This effect also 
leads to a redder continuum slope, as seen in quasars, which is
due to Balmer continuum and Balmer and Fe~\textsc{ii} line emission.
This extreme $89^\circ$ viewing angle cannot represent a typical quasar within a unified model,
but does show that a model such as this can naturally reproduce quasar emission lines
if the emissivity of the wind is increased {\em with respect to the disc continuum}.
In addition, it neatly demonstrates how a stratified outflow can naturally
reproduce the range of ionization states seen in quasars.
% This is because we do not fully solve the C level populations,
% and thus the level populations do not accurately reflect those
% as a full macro-atom with a 
% full collisional rates between forbidden
% or semi-forbidden transitions, as would be required.
% The critical density of the C~\textsc{iii}]~1909\AA\, line 
% is $n_e\sim10^{9.5}~\rm{cm^{-3}}$ \citep{wei1988}, which is higher than much of the 
% outer portion of our wind. 
% We therefore expect a model with these 
% parameters to produce a C~\textsc{iii}]~1909\AA\ line
% with a proper treament. We note, however, the the presence of
% this line in the spectrum of quasars does {\em not} prohibit denser material having a significant contribution to the spectrum.

@@

Despite a number of successes, 
there are a some properties of the synthetic spectra
that are at odds with the observations. First, the ratios of the 
EW of the \la\ and \mgline\ lines
to the EW of \civline\ are much lower than in the composite spectra. 
Similar problems have also been seen in simpler photoionization models for the 
BLR \citep{netzer1990}.
It may be that a larger region of very dense ($n_e\sim10^{10}$cm$^{-3}$) 
material is needed, which could correspond to a disc atmosphere or 
`transition region' 
(see e.g. Murray et al. 1995; Knigge et al. 1997). \nocite{knigge1998} 
While modest changes to geometry may permit this, the initial grid search 
did not find a parameter space in which the \la\ or \mg\ EWs
were significantly higher (see section~\ref{sec:param_sens}). 
Second, we find that EWs increase with inclination 
(see figures~\ref{fig:uvspec} and \ref{fig:sed}; also figure~\ref{fig:lobal}), 
to the extent that, even though significantly denser
models can match the line EWs fairly well at low inclinations, they will then
possess overly strong red wings to the BAL P-Cygni profiles at high inclinations.
The fact that the EW increase in our model are directly related to limb-darkening 
and foreshortening of the continuum. 
This appears to contradict observations, which show remarkably uniform emission
line properties in quasars and BALQSOs \citep{weymann1991,dipompeo2012b}. 
The angular distribution of the disc 
continuum and line emission is clearly crucially important in determining the emergent broad 
line EWs, as suggested by, e.g., the analysis of \cite{risaliti2011}. 
We shall explore this question further in a future study.

@@

The parameter search presented here is by no means exhaustive, and
we may be limited by the specific parameterisation of the outflow 
kinematics we have used. Nevertheless, we suggest that the angular distribution
of both the line and continuum emission is perhaps the crucial 
aspect to understand. With this in mind, obtaining reliable orientation 
indicators appears to be a crucial observational task if we are to
further our understanding of BAL outflows 
and their connection, or lack thereof, to the 
broad line region.

@@

We have carried out MCRT simulations using a simple
prescription for a biconical disc wind, with
the aim of expanding on the work of H13. To do this, we introduced two main
improvements: First, we included a simple treatment of clumping, and second, 
we improved the modelling of recombination lines by treating H and He as
`macro-atoms'. Having selected a fiducial model from an initial simulation grid,
we assessed the viability of such a model for geometric 
unification of quasars, and found the following main points:
\begin{enumerate}
\item Clumping moderates the ionization state
sufficiently to allow for the 
formation of strong UV BALs while agreeing well with the X-ray
properties of luminous AGN and quasars. 
\smallskip
\item A clumpy outflow model naturally 
reproduces the range of ionization states
expected in quasars, due to it's stratified density
and temperature structure. 
LoBAL line profiles are seen at a subset of viewing angles, and Fe~\textsc{ii}
absorption is seen at particularly high inclinations. 
\smallskip
\item The synthetic spectra show \la\ line and weak He~\textsc{ii}~$1640$\AA\ line
as a result of our improved treatment of recombination using macro-atoms. We also see
Balmer emission lines and a Balmer recombination continuum in the optical spectrum, but this
is only really significant at high inclination where to continuum is suppressed.  
\smallskip
\item The higher X-ray luminosity causes a significant 
increase in the strength of the collisionally excited emission
lines produced by the model. 
However, the equivalent width ratios of the emission lines do not match
observations, suggesting that a greater volume of dense ($n_e\sim10^{10}$~cm$^{-3}$)
material may be required.
\smallskip
\item The line EWs in the synthetic spectra increase with inclination.
BAL and non-BAL quasar composites have comparable EWs, so our model
fails to reproduce this behaviour.
 If the BLR emits fairly isotropically then for a 
foreshortened, limb-darkened accretion disc 
it is not possible to achieve line ratios at low inclinations 
that are comparable to those at high inclinations. 
We suggest that understanding the angular distribution of 
line and continuum emission is a crucial question for theoretical models.
% Furthermore, this is conclusion is independent of the assumed 
% BLR geometry and size.
\end{enumerate}
Our work confirms a number of expected outcomes from a geometric unification 
model, and suggests that a simple biconical geometry such as this can come close to 
explaining much of the  phenomenology of quasars. However, our conclusions pose 
some challenges to a picture in which BALQSOs are
explained by an {\em equatorial} wind rising from a classical thin disc, and suggest 
the angular distribution of emission is important to understand if this 
geometry is to be refuted or confirmed. We suggest that obtaining reliable 
observational orientation indicators and 
exploring a wider parameter space of outflow geometries in simulations
are obvious avenues for future work.
